A sheaf--function dictionary for algebraic tori over local fields

### 2013-10-07 shanghai.tex
Addressed the flag in Section 7.

As Proposition~\ref{prop:connected} shows, 
when $G$ is a connected algebraic group over $\Fq$, 
it is appropriate to replace quasicharacter sheaves on $G$ 
with the conceptually simpler category $\QCf(G)$ of finite quasicharacter sheaves on $G$. 
In this context, finite quasicharacter sheaves may also be apprehended as
\emph{certain} $\ell$-adic characters of fundamental group $\pi_1(G,{\bar e})$. 
As explained in \cite{kamgarpour:09a}*{\S~2}, 
the relevant characters of $\pi_1(G,{\bar e})$ are precisely those 
that factor through a particular quotient of $\pi_1(G,{\bar e})$ 
denoted there by $\Pi_\text{disc}(G)$.
%\[\pi_1(G,{\bar e}) \to \Pi_\text{et}(G) \to \Pi_\text{et}(G)_{\Gal(\bFq/\Fq)} = \Pi_\text{disc}(G)\]
This discussion demonstrates that the category of discrete isogenies to $G$ is a Galois category
\cite{grothendieck:SGA1}*{Expos\'e V, \S 4}. Moreover, since $\Pi_\text{disc}(G) \iso G(\Fq)$
\cite{kamgarpour:09a}*{App'x B}, this \David{What does ``this'' refer to} provides an alternate proof that
$\QCiso{G} \iso G(\Fq)^*$ when $G$ is a connected algebraic group over $\Fq$.

@@

\begin{lemma} \label{lem:dual-inv}
Let $X$ be a finitely generated abelian group equipped with an action of $\Weil{}$, and write $\Frob{} = \Frob{}$.  Then
\begin{align*}
 (X^*)_{\Frob{}} &\to (X^{\Frob{}})^* \\
 [f] &\mapsto f|_{X^{\Frob{}}}
\end{align*}
is an isomorphism.
\end{lemma}
\begin{proof}
We can describe $X^{\Frob{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Frob{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Frob{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Frob{}})^*$.
\end{proof}

>>>>

As Proposition~\ref{prop:connected} shows, 
when $G$ is a connected algebraic group over $\Fq$, 
it is appropriate to replace quasicharacter sheaves on $G$ 
with the conceptually simpler category $\QCf(G)$ of finite quasicharacter sheaves on $G$. 
In this context, finite quasicharacter sheaves may also be apprehended as
\emph{certain} $\ell$-adic characters of fundamental group $\pi_1(G,{\bar e})$. 
As explained in \cite{kamgarpour:09a}*{\S~2}, 
the relevant characters of $\pi_1(G,{\bar e})$ are precisely those 
that factor through a particular quotient of $\pi_1(G,{\bar e})$ 
denoted there by $\Pi_\text{disc}(G)$.
%\[\pi_1(G,{\bar e}) \to \Pi_\text{et}(G) \to \Pi_\text{et}(G)_{\Gal(\bFq/\Fq)} = \Pi_\text{disc}(G)\]
This demonstrates that the category of discrete isogenies to $G$ is a Galois category,
in the sense of \cite{grothendieck:SGA1}*{Expos\'e V, \S 4}. 
\Clifton{Is this better?}
Moreover, since $\Pi_\text{disc}(G) \iso G(\Fq)$ \cite{kamgarpour:09a}*{App'x B}, 
the fundamental functor attached to this Galois category can be used to produce an equivalence
between the category of finite quasicharacter sheaves on $G$ and
the category of one-dimensional representations of $G(\Fq)^*$, thus
providing an alternate proof that $\QCiso{G} \iso G(\Fq)^*$ 
when $G$ is a connected algebraic group over $\Fq$.

@@

\Clifton{The finitely generated part of the hypothesis is unnecessary.}
\begin{lemma} \label{lem:dual-inv}
Let $X$ be a finitely generated abelian group equipped with an action of $\Weil{}$, and write $\Frob{} = \Frob{}$.  Then
\begin{align*}
 (X^*)_{\Frob{}} &\to (X^{\Frob{}})^* \\
 [f] &\mapsto f|_{X^{\Frob{}}}
\end{align*}
is an isomorphism.
\end{lemma}
\begin{proof}
We can describe $X^{\Frob{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Frob{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Frob{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Frob{}})^*$.
\end{proof}
